Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2023-01-11 edubots_surveys.Rnw
Revising conclusion

Based on the discussion of this study, we propose a potential process for
introducing technology that would need to be piloted in order to determine its
value and its relationship to improved learning outcomes and higher teacher
staff satisfaction.

Initially, institutions should introduce messaging automation tools that help
with bureaucratic or repetitive tasks, such as answering frequently asked
questions or providing information on class or assignment deadlines.  These
chatbots will pave the way for more complex chatbots that will have a greater
impact on learning outcomes, such as chatbots that help students integrate in
class or identify and address learning problems in individual students or groups
of them. These chatbots should also be accompanied by analytics on student
interaction and possibly some natural language processing (in vernacular
language) to help assess the general mood of the class and how different
materials or external factors affect it.

The introduction of chatbots and chatbot technology will be helped by the fact
that no discernible differences are found between different groups. Even though
chatbots do have some potential for personalization or customization based on
the class material, student, and teaching staff, the general introduction of the
technology can be done in a general way and for all disciplines, teaching
experience, and gender.

Overall, the general feeling that transpires from the survey is that it is
essential for any institution to take into account shareholders' opinions
when introducing chatbots. This is true almost across the board, but in the case
of chatbots (and instant messaging applications) their immediacy and the fact
that they can enter into what we could call the private sphere makes this even
more necessary.

One of the questions in the {\bf second survey}, which asked about the messaging
applications used by tertiary education teachers, opens a new line of inquiry
about what educators {\em perceive} as such, and how it is used.  Namely, the
responses indicated that teachers perceived as messaging application not only
traditional chats or instant messaging app such as WhatsApp or Microsoft Teams,
but also the means provided by their university for communicating with students,
such as a feature of grading applications that emails grades to students. This
suggests that teachers have a need for communication with students, which is
mostly unidirectional, and that it does not matter as much how that need is
met. However, this will need careful consideration, including how it ties with
the automation of the learning experience provided by chatbots.

@@

Finally, the full extent of chatbot technology is not really examined in these
surveys. They can be connected to natural language processing engines with
sentiment analysis as well as other analytics. Examining the mood of the class
such as responses to new materials, assignments, exams or external events will
help any student-centered teaching strategy, which will also help students (and
teachers) achieve their learning objectives. This line of work, however, would
require the introduction of some pilot study combined with initial opinion
assessments from students.

>>>>

Based on the discussion of this study, we propose here a potential process for
introducing technology that would need to be piloted in order to determine its
value and its relationship to improved learning outcomes and higher teacher
staff satisfaction.

Initially, institutions should introduce messaging automation tools attached to
some, or several, messaging platforms that are already popular among the
university community, that would help with bureaucratic or repetitive tasks,
such as answering frequently asked questions or providing information on class
or assignment deadlines.  These chatbots will pave the way for more complex ones
that will have a greater impact on learning outcomes, such as chatbots that help
students integrate with other students in class or identify and address learning
problems in individual students or groups of them. These chatbots should also be
accompanied by analytics on student interaction and possibly some natural
language processing (in vernacular language) to help assess the general mood of
the class and how different materials or external factors affect it. These tools
would have to be attached to a learning analytics platform, which, in turn,
would be part of an institutional learning management system.

The introduction of chatbots and chatbot technology will be helped by the fact
that no discernible differences are found in the survey between different
demographic groups. Even though chatbots do have some potential for
personalization or customization based on the class material, student, and
teaching staff, the general introduction of the technology can be done in a
general way and for all disciplines, types or degrees, teaching experience, and
gender.

Overall, the general feeling that transpires from the survey is that it is
essential for any institution to take into account stakeholders' opinions when
introducing chatbots. This is true almost across the board for any new
technology, but in the case of chatbots (and instant messaging applications)
their immediacy and the fact that they can enter into what we could call the
private sphere makes this even more necessary.

One of the questions in the {\bf second survey}, which asked about the messaging
applications used by tertiary education teachers, opens a new line of inquiry
about what educators {\em perceive} as such, and how it is used.  Namely, the
responses indicated that teachers considered messaging application not only
traditional chats or instant messaging app such as WhatsApp or Microsoft Teams,
but also any means provided by their university for communicating with students,
such as a feature of the cmapus-wide learning management system that emails
grades to students. This suggests that teachers have a need for communication
with students, which is mostly unidirectional, and that it does not matter as
much how that need is met. However, this will need careful consideration,
including how it ties with the automation of the learning experience provided by
chatbots.

@@

Finally, the full extent of chatbot technology is not really examined in these
surveys. They can be connected to natural language processing engines with
sentiment analysis as well as other analytics. Examining the mood of the class
such as responses to new materials, assignments, exams or external events will
help any student-centered teaching strategy, which will also help students (and
teachers) achieve their learning objectives. This line of work, however, would
require the introduction of some pilot study combined with initial opinion
assessments from students and is thus left for the future.
